Speed and strength of epidemic intervention

### 2019-11-13 doc/speedstrength.tex
writing is hard

The study of measles dynamics provides a key example for which the characterization of an epidemic via the reproduction number has been successful.
One of the earliest estimates of the measles $\RR_0$ comes from \cite{anderson1982directly} nearly 40 years ago, who gave a range of $14 - 18$.
This, combined with the mean age of infection (4.6 years) and the mean age of vaccination (2.2 years), allowed them to predict that approximately 96\% of the new birth cohort had to be vaccinated to eradicate measles.

Like epidemic strength, epidemic speed measures how disease spreads in the population. 
\swp{Ebola}

>>>>

The study of measles dynamics provides a key example for which the characterization of an epidemic via the basic reproduction number has been successfully applied in controlling the disease.
One of the earliest estimates of the measles $\RR_0$ comes from \cite{anderson1982directly} nearly 40 years ago, who gave a range of $14 - 18$.
This, combined with the mean age of infection (4.6 years) and the mean age of vaccination (2.2 years), allowed them to predict that approximately 96\% of the new birth cohort had to be vaccinated to eradicate measles.
Since then, measles control policies have revolved around targetting this critical vaccination threshold \citep{venczel2003measles, wolfson2007has, takahashi2015reduced}.
\swp{More citations later}

Like epidemic strength, epidemic speed measures how disease spreads in a population.
Despite the prevailing focus on the basic reproduction number in literature \swp{CITE}, 
relying on epidemic speed can be sometimes more effective in characterizing the course of an epidemic and evoluating intervention strategies.
In particular, the exponential growth rate can be directly measured from incidence data even early in an epidemic \citep{chowell2003sars, mills2004transmissibility, nishiura2009transmission, nishiura2010pros, ma2014estimating}, whereas the basic reproduction number can be difficult to estimate \citep{weitz2015modeling};
in these scenarios, intervention strategies should seek to reduce the epidemic speed.
The West African Ebola outbreak in 2014 provides a key example for this perspective: 
many researchers tried to estimate the basic reproduction number but the major intervention strategies involved slowing down the epidemic by identifying infected individuals and isolating them \citep{pandey2014strategies}.
